Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-03-23 bbbq_article.tex
Processed feedback 'Measuring TMH epitopes', thanks @fransbianchi

% \paragraph{Data sets for TMH epitopes}
A human, viral and bacterial reference proteome was used to determine the percentages of epitopes overlapping with TMHs. The human reference proteome has UniProt ID UP000005640_9606, which is a representative set of the full proteome and also used in our previous study (\cite{bianchi2017}).  SARS-CoV-2 with reference proteome UniProt ID UP000464024 was used as a viral proteome and Mycobacterium tuberculosis (MTb) with UniProt ID UP000001584 was used . These pathogens were chosen based on their impact on the human population(references).

@@

To determine if epitopes derived from TMHs are presented we used epitopes identified in \frans{Welke studies is dit dan van SARS-COV2 en MTB ?}
for MHC-I (\cite{schellens2015comprehensive}) 
and MHC-II (\cite{bergseng2015different})peptide elution studies.
For each of these peptides, the human representative reference proteome was searched for its origin\frans{wat bedoel je?}. We filtered for epitopes that were uniquely present in the proteome and correlated these for overlap with predicted transmembrane regions in the proteome. 
In this analysis, both TMHMM and PureseqTM (see section ’Prediction software used’) were used to predict the transmembrane topology. 
The full analysis can be found
at \url{https://github.com/richelbilderbeek/bbbq_article_issue_157}.

@@

To determine the percentages of MHC-I and MHC-II epitopes overlapping with TMHs, we applied a similar analysis strategy as described in  \cite{bianchi2017}.
To summarize: from a proteome all possible 9-mers are derived. For each
of these peptides, it was determined if it was part of a 
TMH (at least overlapping with 1 residue), 
and if the peptide bound to an MHC-I haplotype.
For MHC-II, 14-mers were used, as these are the most frequently occurring
epitope length (\cite{bergseng2015different}).

@@

The 13 MHC-I haplotypes used in this study are the same as 
used in our previous study (\cite{bianchi2017}).
The MHC-II haplotypes used additionally were selected 
to occur with a phenotypic frequency of at least 14\% in
the human population (\cite{greenbaum2011functional}),
resulting in 21 haplotypes.
When using an MHC-II haplotype, instead of using 9-mers, 14-mers were
used, as these are the most common MHC-II epitope size,
as found in the elution study (\cite{bergseng2015different}).\frans{what percentage is 9 mer or 14 mer?}

>>>>

@@

To determine if epitopes derived from TMHs are presented 
we used two elution studies, which both have sequenced
epitopes found in naturally occurring human immune cells,
for MHC-I (\cite{schellens2015comprehensive}) 
and MHC-II (\cite{bergseng2015different}).
For each of the sequenced epitopes,
we found its possible location(s) in 
a human representative reference proteome,
with UniProt ID UP000005640_9606.
For the epitopes that were found exactly once in the proteome,
we predicted the topology of the proteins in which the epitopes are found.
We used both TMHMM and PureseqTM to do these predictions. 
From this topology, we could determine if the epitope
overlapped with a transmembrane regions.

The full analysis can be found
at \url{https://github.com/richelbilderbeek/bbbq_article_issue_157}.

@@

To determine how frequently epitopes overlapping with TMHs are presented,
we applied a similar analysis strategy as described in \cite{bianchi2017}.
To investigate the generality of this, we applied this analysis
for several haplotypes of both MHC-I and MHC-II, 
as well as for a human, viral and bacterial proteome.
To summarize, for each proteome, 
all possible 9-mers (for MHC-I) or 14-mers (MHC-II) are derived. 
For each of these peptides, we determined if it overlapped with a 
TMH and if it was a binder for each haplotype.

For MHC-I, 9-mers were used, as this is the length used in our earlier
study (\cite{bianchi2017}). For MHC-II, 14-mers were used, 
as these are the most frequently occurring
epitope length (\cite{bergseng2015different}).
We used a human (UniProt ID UP000005640_9606), 
viral (SARS-CoV-2, UniProt ID UP000464024) 
and bacterial (Mycobacterium tuberculosis (MTb), UniProt ID UP000001584) 
reference proteome. The pathogens were chosen based 
on their impact on the human 
population \frans{Add references here}.
We used TMHMM  \richel{ref} to predict the topology 
of the proteins within the proteome.
To predict the affinity of an epitope to a certain MHC molecule,
we used \verb;EpitopePrediction; (\cite{bianchi2017}) for MHC-I 
and \verb;MHCnuggets; \richel{ref} for MHC-II.
The 13 MHC-I haplotypes used in this study are the same as 
used in our previous study (\cite{bianchi2017}).
The MHC-II haplotypes used additionally were selected 
to occur with a phenotypic frequency of at least 14\% in
the human population (\cite{greenbaum2011functional}),
resulting in 21 haplotypes.

@@
